Fig5

### 2020-03-27 analysis/figure-3.Rmd
reorder figures

---
title: "Fig3"
output: md_document
---

# Figure 3

```{r setup}
library(tidyverse)
library(cowplot)
library(ggimage)
library(apaTables)
library(factoextra)

source("../R/themes.R") 
source("../R/functions.R")
source("../R/wrangledata.R")

knitr::opts_chunk$set(fig.path = '../figures/',message=F, warning=FALSE)
```

## Data: PCA, hormones, PRL expression

```{r PRLfoo}
# pca
charpca <- subsetmakepca("pituitary", charlevels, sexlevels)	
charfviz <- makefvizdf("pituitary", charlevels, sexlevels)

# hormones
prolactin <- read_csv("../results/07_hormones.csv") %>%
    filter(study == "characterization", hormone %in% c("prolactin"))  %>% 
    droplevels() 
prolactin$treatment <- factor(prolactin$treatment, levels = alllevels)

# PRL vsd

@@

PRLvsd <- vsd_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name")  %>% 
  dplyr::rename("gene" = "X1") %>% 
  pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
               names_to = "samples", values_to = "counts") %>%
  filter(gene == "PRL")  %>%
  dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1)) %>%
  dplyr::mutate(sextissue = sapply(strsplit(sextissue, '../results/DEseq2/'), "[", 2)) %>%
  dplyr::mutate(sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
                tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
                treatment = sapply(strsplit(samples, '\\_'), "[", 4)) %>%
  dplyr::mutate(treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1)) %>%
  dplyr::select(sex, tissue, treatment, gene, samples, counts) %>%
  drop_na()
PRLvsd$treatment <- factor(PRLvsd$treatment, levels = alllevels)

PRLhyp <- PRLvsd %>% filter(tissue == "hypothalamus") 
PRLpit <- PRLvsd %>% filter(tissue == "pituitary")
PRLgon <- PRLvsd %>% filter(tissue == "gonad")
```

```{r determinePRLhiglo, fig.height = 6, fig.width=3}
PRLvsd2 <- PRLvsd %>%
  filter(tissue == "pituitary")
head(PRLvsd2)

PRLvsd2 %>%
  group_by(sex) %>%
  summarize(median = median(counts))

PRLvsd3 <- PRLvsd2 %>%
  mutate(hiloPRL = ifelse(counts >= 18, "hi", "lo"))  %>%
  drop_na()
PRLvsd3$hiloPRL <- factor(PRLvsd3$hiloPRL, levels = c("lo", "hi"))

PRLvsd3 %>%
  group_by(sex, tissue, hiloPRL) %>%
  summarize(n = n())

@@

## Internal versus external

```{r hypothesis}
DEG_path <- "../results/DEseq2/hypothesis/"   # path to the data
DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
DEG_pathfiles <- paste0(DEG_path, DEG_files)
#DEG_files

allDEG2 <- DEG_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name") %>% 
  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/hypothesis/'), "[", 2))  %>% 
  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
  mutate(comparison = paste(down,up, sep = "_")) %>%
  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) 
head(allDEG2)

allDEG2$tissue <- factor(allDEG2$tissue, levels = tissuelevel)

allDEG2$comparison <- factor(allDEG2$comparison, levels = c("eggs_chicks", "lo_hi"))
allDEG2 <- allDEG2 %>% mutate(comparison = fct_recode(comparison, "lo vs. hi PRL   " = "lo_hi",
                                                      "eggs vs. chicks" = "eggs_chicks"))
allDEG2$direction <- factor(allDEG2$direction, levels = c("eggs", "chicks", "lo", "hi"))

```{r fig3, fig.width=7, fig.height=7}

expdesign <- png::readPNG("../figures/images/fig_fig3a.png")
expdesign <- ggdraw() +  draw_image(expdesign, scale = 1)

b <- plotcolorfulpcs(charpca,charpca$treatment, allcolors) + labs(subtitle = " ") +
  theme(legend.position = "none", 
        legend.direction = "horizontal", 
        legend.key.size = unit(0.5, 'lines')) + 
  guides(color = FALSE) +
  labs(subtitle = "pituitary")

c <- plotprolactin(PRLpit, PRLpit$counts, "PRL", "pituitary") + 
  theme(legend.position = "none", 
         axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(face = "italic"))

d <- plotprolactin(prolactin, prolactin$plasma_conc, "prolactin (ng/mL)", "blood") +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))

bcd <- plot_grid(b,c,d, labels = c("b", "c", "d"), ncol = 1, label_size = 8, rel_heights = c(1,1,1.3))

makenewbargraph <- function(whichtissue, whichsex,  whichcomparison, lowlim, higherlim){
  p <- allDEG2 %>%
    filter(tissue == whichtissue,
           comparison == whichcomparison,
           sex == whichsex) %>%
    ggplot(aes(x = comparison,  fill = direction)) +
    geom_bar(position = "dodge", drop = FALSE) +
    theme_B3() +
    theme(legend.position = "none")  +
    guides(fill = guide_legend(nrow = 1)) +
    labs( y = whichtissue) +
  geom_text(stat='count', aes(label=..count..), vjust =-0.5, 
           position = position_dodge(width = 1),
           size = 2, color = "black")  + 
      ylim(lowlim, higherlim) +
  scale_fill_manual(values = allcolors, name = "higher in")   + 
    theme(axis.text.x = element_blank()) 
  return(p)
}

b11 <- makenewbargraph("hypothalamus", "female","eggs vs. chicks", 0, 2500) +  labs(subtitle = "females", x = NULL, y = " ") + 
  theme( axis.text = element_blank())
b21 <- makenewbargraph("pituitary", "female","eggs vs. chicks", 0, 2500)  + labs(x = NULL, y = " ") + 
  theme(axis.text = element_blank())
b31 <- makenewbargraph("gonad", "female", "eggs vs. chicks", 0, 2500) + labs(x = "eggs vs. chicks", y = " ")    + 
  theme(axis.text = element_blank())
b112131 <- plot_grid(b11,b21,b31, nrow = 3, rel_heights = c(1.1,1,1.1))

b12 <- makenewbargraph("hypothalamus", "male",  "eggs vs. chicks", 0, 2500) + labs(subtitle = "males", x = NULL)+ 
  theme(axis.title.y = element_blank(), axis.text = element_blank())
b22 <- makenewbargraph("pituitary", "male", "eggs vs. chicks", 0, 2500) + labs(x = NULL)+ 
  theme(axis.title.y = element_blank(), axis.text = element_blank())
b32 <- makenewbargraph("gonad", "male", "eggs vs. chicks", 0, 2500) + labs(x = "eggs vs. chicks")   + 
  theme(axis.title.y = element_blank(), axis.text = element_blank())
b122232 <- plot_grid(b12, b22,b32, nrow = 3, rel_heights = c(1.1,1,1.1))

c11 <- makenewbargraph("hypothalamus", "female", "lo vs. hi PRL   ", 0, 2500) +  labs(subtitle = "females", x = NULL, y = "hypothalamus DEGs") + 
  theme( axis.text.x = element_blank()) 
c21 <- makenewbargraph("pituitary", "female", "lo vs. hi PRL   ", 0, 2500)  + labs(x = NULL, y = "pituitary DEGs")+ 
  theme( axis.text.x = element_blank())
c31 <- makenewbargraph("gonad", "female", "lo vs. hi PRL   ", 0, 2500) + labs(x = "lo vs. hi PRL", y = "gonad DEGs")  + 
  theme(axis.text.x = element_blank()) 
c112131 <- plot_grid(c11,c21,c31, nrow = 3, rel_heights = c(1.1,1,1.1))

c12 <- makenewbargraph("hypothalamus", "male",  "lo vs. hi PRL   ", 0, 2500) + labs(subtitle = "males", x = NULL)+ 
  theme(axis.title.y = element_blank(), axis.text = element_blank())
c22 <- makenewbargraph("pituitary", "male", "lo vs. hi PRL   ", 0, 2500) + labs(x = NULL)+ 
  theme(axis.title.y = element_blank(), axis.text = element_blank())
c32 <- makenewbargraph("gonad", "male", "lo vs. hi PRL   ", 0, 2500) + labs(x = "lo vs. hi PRL")   + 
  theme(axis.title.y = element_blank(), axis.text = element_blank())
c122232 <- plot_grid(c12, c22,c32, nrow = 3, rel_heights = c(1.1,1,1.1))

hypothesisbars <-  plot_grid(c112131, c122232, b112131, b122232, nrow = 1, rel_widths = c(1.3, 1, 1.1, 1),
          labels = c("f", " ", "g", " "), label_size = 8)

dataplots <- plot_grid(bcd, hypothesisbars)

fig3 <- plot_grid(expdesign, dataplots, rel_heights = c(0.25,1), nrow = 2, labels = c("a"), label_size = 8)
fig3

```{r savefiles}
#write.csv(PRLvsd3,"../results/PRLvsd.csv", row.names = F)

>>>>

---
title: "Fig3"
author: "Rayna M Harris"
date: "3/26/2020"
output: md_document
---

# all things box plots

## candidate genes

```{r}
library(tidyverse)
library(ggtext)
library(cowplot)
library(ggpubr)

source("../R/themes.R")

knitr::opts_chunk$set(echo = TRUE, message = F, fig.path = "../figures/")

candidategenes <- c("OXT", "AVP", "GNRH1", "GNRHR", "CGNRH-R",
                    "AR", "POMC", "AGRP", 
                       "CRH", "AVPR1A", "AVPR1B", "AVPR2","VIP",
                       "CYP19A1", "DRD1", "DRD2", "PRL", "PRLR", "SOX9", 
                    "ESR1","ESR2", "LBH", "CDK1", "BRCA1",
                    "PTEN", "CREBBP", "FOS", "JUN", "EGR1",
                     "BDNF", "GRM2","GRIA1",
                    "KCNJ5", "CISH", "PTGER3", "CEBPD", "ZBTB16", 
                    "DIO3", "DIO2", "DIO1",

                     "VIPR1", "VIPR2", "NPY", "LH", "FSH", "FSHB",
                    "JAK2", "HCRT", "TRH", "TSHB",
                    "MC3R", "MC4R", "MC5R",  "NR3C1", "NR3C2",
                    "STAT5B","NPVF")

candidategenesslim <- c("OXT", "AVP", "GNRH1", "GNRHR", 
                    "AR",  "CYP19A1", 
                     "AVPR1A", "AVPR1B", "AVPR2","VIP",
                  "DRD1", "DRD2", 
                  "PRL", "PRLR",  
                    "ESR1","ESR2", "LBH",

                    "FOS", "JUN", "EGR1", "BDNF"
                      )

## variance stabilized gene expression  (vsd)

```{r }
geneids <- read_csv("../metadata/00_geneinfo.csv")

@@

allvsd <- vsd_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name")  %>% 
  dplyr::rename("gene" = "X1") %>% 
  pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
               names_to = "samples", values_to = "counts")

getcandidatevsd <- function(whichgenes, whichtissue, whichsex){
  candidates  <- allvsd %>%
    filter(gene %in% whichgenes) %>%
    dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1)) %>%
    dplyr::mutate(sextissue = sapply(strsplit(sextissue, '../results/DEseq2/'), "[", 2)) %>%
    dplyr::mutate(sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
                tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
                treatment = sapply(strsplit(samples, '\\_'), "[", 4)) %>%
    dplyr::mutate(treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1)) %>%
    dplyr::select(sex, tissue, treatment, gene, samples, counts) %>%
    filter(tissue == whichtissue, sex %in% whichsex)  %>%
    drop_na()
  candidates$treatment <- factor(candidates$treatment, levels = alllevels)
  return(candidates)
}

hypvsd <- getcandidatevsd(candidategenesslim, "hypothalamus", sexlevels)
pitvsd <- getcandidatevsd(candidategenesslim, "pituitary", sexlevels)
gonvsd <- getcandidatevsd(candidategenesslim, "gonad", sexlevels)
head(hypvsd)
head(pitvsd)
head(gonvsd)

candidatevsd <- rbind(hypvsd, pitvsd)
candidatevsd <- rbind(candidatevsd, gonvsd)

unique(candidatevsd$gene)
```

## Figs

```{r fig3, fig.width=7, fig.height=7}
makeboxplots <- function(df, whichgene, mysubtitle, whichsex, whichtimepoint){
  p <- df %>%
    filter(treatment %in% charlevels,
           gene %in% whichgene,
           sex %in% whichsex) %>%
    filter(treatment %in% whichtimepoint) %>%
    ggplot(aes(x = treatment, y = counts, fill = treatment, color = sex)) +
    geom_boxplot(outlier.shape = NA) + 
    geom_jitter(size = 0.5, width = 0.1) +
    facet_wrap(~gene, scales = "free_y", nrow = 1) +
    theme_B3() +
    scale_fill_manual(values = allcolors) +
    scale_color_manual(values = allcolors) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none") +
    labs(y = "expression" , x = "Parental stages", subtitle = mysubtitle) +
    theme(strip.text = element_text(face = "italic"))
  return(p)
}

## hyp 
a <- makeboxplots(hypvsd, c("BDNF","CYP19A1", "DRD1", "EGR1"), 
                  "Female hypothalamus", "female", c("hatch", "n5"))  +
  theme(axis.title.x = element_blank())

c <- makeboxplots(pitvsd, c("ESR1","GNRHR"), "Female pituitary", "female", c("bldg", "lay", "inc.d3"))   +
  theme(axis.title.x = element_blank())
d <- makeboxplots(pitvsd, c("LBH", "PRL" ), " ", "female", c("inc.d9", "inc.d17", "hatch", "n5"))  +
  theme(axis.title  = element_blank())
e <- makeboxplots(pitvsd, c( "AVPR2"), " ", "female", c( "hatch", "n5"))  +
  theme(axis.title  = element_blank())

cde <- plot_grid(c,d, e, rel_widths = c(2.5,2.5,1), nrow = 1)

g <- makeboxplots(gonvsd, c( "AVPR1A"), "Female gonads", "female", c( "lay", "inc.d3", "inc.d9"))  +
  theme(axis.title.x = element_blank())
h <- makeboxplots(gonvsd, c( "EGR1", "FOS", "PRLR"), " ", "female", c( "lay", "inc.d3"))  +
  theme(axis.title = element_blank())

gh <- plot_grid(g,h,nrow = 1, rel_widths = c(1,2.5))

b <- makeboxplots(hypvsd, c("AR"),"Male hypothalamic expression", "male", 
                  c( "inc.d9", "inc.d17"))  + labs(caption = " ")
f <- makeboxplots(pitvsd, c("LBH", "PRL"), "Male pituitary expression", "male", 
                  c("inc.d9", "inc.d17")) + 
  theme(axis.title.y  = element_blank(),
        plot.caption = element_text(face = "italic")) + 
  labs(caption = "candidate genes: AR, AVP, AVPR1A, AVPR1B, AVPR2, BDNF, CYP19A1, DRD1, EGR1, ESR1, ESR2, FOS, GNRH1, GNRHR, JUN, LBH, OXT, PRL, PRLR, VIP")

bf <- plot_grid(b,f,nrow = 1, rel_widths = c(1,2))

plot_grid(a, cde, gh, bf, nrow = 4, rel_heights = c(1,1,1,1.2), labels = "auto", label_size = 8)

@@

```{r writefiles}
write.csv(candidatevsd, "../../musicalgenes/data/candidatecounts.csv")
```

### 2020-03-27 analysis/figure-5.Rmd
reorder figures

```{r setup, include=FALSE}
library(tidyverse)
library(cowplot)
source("../R/themes.R")

knitr::opts_chunk$set(echo = TRUE, message = F, fig.path = "../figures/")
```

## DEGs

```{r fig2, fig.width=7}

DEG_path <- "../results/DEseq2/manip/"   # path to the data
DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
DEG_pathfiles <- paste0(DEG_path, DEG_files)
DEG_files

allDEG <- DEG_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name") %>% 
  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/manip/'), "[", 2))  %>% 
  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
  mutate(comparison = paste(down,up, sep = "_")) %>%
  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) 
head(allDEG)

comparisonlevels <- c( "inc.d3_m.inc.d3" ,  "inc.d9_m.inc.d9", 
                        "inc.d17_m.inc.d17", "m.n2_hatch" ,
                        "inc.d9_m.inc.d8", "inc.d17_prolong", "hatch_extend"  )

allDEG$tissue <- factor(allDEG$tissue , levels = tissuelevel)
allDEG$comparison <- factor(allDEG$comparison , levels = comparisonlevels)
allDEG$direction <- factor(allDEG$direction, levels = manipulationsandcontrols)

allDEG %>%
  group_by(sex, tissue, comparison) %>%
  summarize(n = n())

allDEG

allDEG %>%
ggplot(aes(x = comparison,  fill = direction)) +
  geom_bar(position = "dodge") +
  facet_grid(tissue~sex) +
  theme_B3() +
  theme(#axis.text.x = element_text(angle = 45, hjust = 1),
         axis.text.x = element_blank(),
        legend.position = "none",
        strip.placement = "right")  +
  guides(fill = guide_legend(nrow = 1)) +
  labs(x = "Parental stage-specific manipulations", y = NULL) +
  scale_fill_manual(values = allcolors,
                       name = " ", drop = FALSE) +
  scale_color_manual(values = sexcolors) +
  geom_text(stat='count', aes(label=..count..), vjust =-0.5, 
            position = position_dodge(width = 1),
            size = 2.5, color = "black")

@@

>>>>

```{r setup, include=FALSE}
library(tidyverse)
library(cowplot)
library(Rtsne)

source("../R/themes.R")
source("../R/functions.R")
source("../R/wrangledata.R")

knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/')
```

# Figure 1. Experimental design, tSNE analysis, PCA, and prolactin

```{r tSNE-df}

# prep for tsne

hyptsne <- subsetmaketsne("hypothalamus", allmaniplevels, sexlevels)
pittsne <- subsetmaketsne("pituitary", allmaniplevels, sexlevels)
gontsne <- subsetmaketsne("gonads", allmaniplevels, sexlevels)

addgroupings <- function(df){

  df <- df %>% mutate(hiloPRL = fct_collapse(treatment, 
                                  lo = c("m.inc.d8", "m.inc.d3", "m.inc.d9", "inc.d3", "inc.d9", "lay", "bldg"),
                                  hi = c("inc.d17",   "m.inc.d17", "prolong" ,  "hatch" ,  "m.n2", "extend", "n5")),
                extint = fct_collapse(treatment, 
                                  eggs = c("inc.d3", "inc.d9", "inc.d17", "prolong", "lay"),
                                  chicks = c( "m.inc.d8", "hatch", "extend", "n5"),
                                  loss = c("m.inc.d3", "m.inc.d9", "m.n2", "m.inc.d17", "bldg")))
  df$extint <- factor(df$extint, levels = levelsextint)
  return(df)
}

hyptsne <- addgroupings(hyptsne)
pittsne <- addgroupings(pittsne)
gontsne <- addgroupings(gontsne)

@@

## make figure

```{r fig5, fig.width=7, fig.height=7}

expdesign2 <- png::readPNG("../figures/images/fig_fig4a.png")
expdesign2 <- ggdraw() +  draw_image(expdesign2, scale = 1)

b <- plottsneelipsev2(hyptsne, hyptsne$treatment, allcolors) + labs(subtitle = "hypothalamus ~ parental stage", x = NULL)  + facet_wrap(~sex, scales = "free")
c <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors ) + labs(subtitle = "pituitary ~ parental stage", x = NULL) + facet_wrap(~sex, scales = "free") 
d <- plottsneelipsev2(gontsne, gontsne$treatment, allcolors ) + labs(subtitle = "gonads ~ parental stage")  + facet_wrap(~sex, scales = "free")

bcd <- plot_grid(b,c,d, nrow = 3, labels = c("b"), label_size = 8, rel_heights = c(1,1,1))

e <- plottsneelipsev2(hyptsne, hyptsne$hiloPRL, allcolors ) + labs(subtitle = "hypothalamus ~ internal physiology", x = NULL) + facet_wrap(~sex, scales = "free")  
f <- plottsneelipsev2(pittsne, pittsne$hiloPRL, allcolors ) + labs(subtitle = "pituitary ~ internal physiology", x = NULL) + facet_wrap(~sex, scales = "free")   
g <- plottsneelipsev2(gontsne, gontsne$hiloPRL, allcolors ) + labs(subtitle = "gonads ~ internal physiology" ) + facet_wrap(~sex, scales = "free")

efg <- plot_grid(e,f,g, nrow = 3, labels = c("d"), label_size = 8, rel_heights = c(1,1,1))

h <- plottsneelipsev2(hyptsne, hyptsne$extint, allcolors ) + labs(subtitle = "hypothalamus ~ external enviornment", x = NULL) + facet_wrap(~sex, scales = "free")  
i <- plottsneelipsev2(pittsne, pittsne$extint, allcolors ) + labs(subtitle = "pituitary ~ external enviornment", x = NULL) + facet_wrap(~sex, scales = "free")   
j <- plottsneelipsev2(gontsne, gontsne$extint, allcolors ) + labs(subtitle = "gonads ~ external enviornment" ) + facet_wrap(~sex, scales = "free")

hij <- plot_grid(h,i,j, nrow = 3, labels = c("c"), label_size = 8, rel_heights = c(1,1,1))

bcdhijefg <- plot_grid(bcd, hij,  efg, nrow = 1)

fig5 <- plot_grid(expdesign2, bcdhijefg, nrow = 2, labels = c("a"), label_size = 8, rel_heights = c(0.25,1))
fig5
